---
title: "State of Disproportionality"
author: "Shannon Coulter"
date: "November 23, 2021"
output:
  html_document:
    theme: readable
    highlight: espresso
  word_document: default
  pdf_document: default
subtitle: Understanding Policies and Procedures across Organizations
editor_options:
  markdown:
    wrap: 72
---

```{r rmarkdown-setup, include=FALSE}
#https://stats.andrewheiss.com/ngo-crackdowns-philanthropy/02_analysis.html
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(warning = FALSE)
options(knitr.kable.NA = "")
options(kableExtra.latex.load_packages = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(here)
library(gtsummary)
library(gt)
```

## State of Disproportionality {.tabset .tabset-fade .tabset-pills}

### Survey Overview

We administered the State of Disproportionality Survey to special
education directors across California to gather information about
disproportionality and the work districts engage in to reduce it. The
South County SELPA, one of four SELPA Content Leads for the State of
California, designed this survey as part of a state grant to address
disproportionality. We designed the survey to capture the "state" of
disproportionality in each of California's 144 SELPA regions by asking
questions about experiences regarding special education data, culture,
general education processes and policies, training, system of support,
resources, and others. <br><br> We sent the survey to approximately 940
school districts and received 320 responses for a response rate of 34%.
Several factors affected the lower response rate. First, many districts
do not receive disproportionality results because of the size of the
school district and student populations. These districts elected not to
take the survey. Second, some districts do not have disproportionality
findings according to the state improvement plan because there are no
differences in proportions of students in special education, such is the
case in districts where the population is largely homogeneous. This is
another reason why some districts elected not to take the survey. Third,
because disproportionality is a shifting metric each year (i.e., a
district may not have a finding one year and have one the next)
determining the population to survey is challenging. Therefore, our
response rate may be lower simply because we sent the survey to a
greater number of districts who would have actually completed it.
<br><br> We share the results using contingency tables where rows are
the survey items and columns are organizations supporting districts with
disproportionality work. Below we arrayed the participants by the
organizations supporting them with disproportionality work. Most
participants worked with the System Improvement Leads (sils), followed
by SELPA content leads, SPP-TAP, WestEd, System Improvement Leads,
Orange County, ED&D, Cal-TAC, and CDE's Seeds of Partnership. About 23
participants completing the survey worked with ED&D.

```{r child="sections/demographics.Rmd"}
```

### Disproportionality Knowledge

We divided the survey into four sections. This section focuses on
disproportionality knowledge. Participants were asked a series of
questions to gauge their knowledge level on disproportionality. We
arrayed the results in contingency tables depicting the
statement/responses and the organization for which the survey
participants received support. We displayed the number of participants
completing the survey under each organization's abbreviated name along
with the frequency of responses for each category of the survey item.

```{r child="sections/knowledge.Rmd"}
```

### General Education Practices

This section highlights general education processes and procedures
leading to a special education referral. These questions ask
participants the degree to which these processes are in place in their
district and the extent to which the district implements these processes
well. We arrayed the results by statement and organization and the
number of participants completing the survey.

```{r child="sections/general_education.Rmd"}
```

### Mutli-tiered System of Supports

The following section highlights MTSS processes and procedures present
in the school district. These questions ask whether the district has
these processes and procedures in place and the extent to which the
district implements these processes and procedures well. We arrayed the
results by statement and organization with the number of participants
completing the survey.

```{r child="sections/mtss.Rmd"}
```

### Disproportionality Practices, Resources, and Supports

The following section of the survey highlights the section on
disproportionality practices, resources, processes and procedures, and
supports. This section asks respondents to respond to twenty-eight
questions related to their districts organizational capacity, resources,
practices, and policies regarding disproportionality. Similar to
previous sections, we arrayed the results by statement and organization
with the number of participants completing the survey listed under the
organization's name.

```{r child="sections/dispro_practices.Rmd"}
```

### Findings and Conclusions

We administered a survey on the state of disproportionality to special
education administrator's across California. We recieved 320 reponses
and arrayed these responses by the organizations supporting districts
with disproportionality-related services. Organizations reflected in the
report include ED&D, System Improvement Leads (SILS), Supporting
Inclusive Practices (SIP), and SPP-TAP. <br><br> **The System
Improvement Leads registered the highest percentages of favorable
responses.**<br> In 21 of the 48 survey items (or 44% of the survey),
participants supported by the System Improvement Leads rated items
qualitatively better than participants supported by other organizations.
For example, when asked: "Does your district gather data on the ethnic
and racial breakdowns of children being referred to special education?",
78% of respondents supported by the SILS indicated "Yes" to that
question, which is a preferred or recommended behavior for a district
addressing disproportionality. It is important to note that differences
between responses aggregated to an organizational level are often quite
small, sometimes only a percentage point difference. Another important
consideration is that we do not know if the systems supported by the
SILS would have rated these items higher prior to working with the SILS.
The survey represents a "snapshot" of how well equipped districts are to
address disproportionality (arrayed by the organizations supporting
them). It appears based on survey responses that districts supported by
SILS are best situated to reduce disproportionality in the future.
<br><br> **The support ED&D and SPP-TAP provide have a similar
percentage of favorable responses.**<br> Participants supported by ED&D
rated 11 items qualitatively better (23%) than respondents reporting on
behalf of other organizations, which is the same rate as respondents
reporting on behalf of SPP-TAP. Participants from these two
organizations appear equally well equipped to reduce disporportionality
in the future. <br><br>

*ED&D Strengths and Weaknesses*

In terms of the major sections of the survey: Disproportionality
Knowledge, General Education Practices, MTSS, and Disproportionality
Practices, Resources, and Supports, ED&D performs best in the last
section around disproportionality practices, resources, and supports.
<br><br> **ED&D outperforms all other support organizations in
disproportionality practices, resources, and supports.**<br> For
example, ED&D respondents indicated they believed their district's risk
ratio scores would decline in the coming year, their districts are
better equipped to tackle disproportionality, their districts have
enacted policies to prevent disproportionality, and their districts have
specific goals to address disproportionality. ED&D received the highest
percentage of preferred/favorable responses to the questions related to
disproportionality practices, resources, and supports. SPP-TAP performed
similarly to ED&D in this section of the survey, being edged out by a
few questions. SPP-TAP received the highest percentage of
preferred/favorable responses to 38% of the survey items in this
section.<br><br> **ED&D must consider how its current services align and
integrate with other drivers of disproportionate outcomes.**<br> Many
systems influence and affect special education outcomes. Support
organizations (such as ED&D), regardless of their core services, must be
mindful of the impact each system has on disproportionality. If a car
doesn't start and a mechanic focuses all of his/her energy repairing the
electrical system, the car is no more functional if its transmission
does not adjust the gearing in the car. Therefore, ED&D must consider
how its services align and integrate with other drivers of
disproportionate outcomes. For example, districts supported by ED&D
indicated lower percentages of preferred/favorable responses compared to
other support organizations in relation to convening SSTs and
implementing a process to review evidence prior to making a special
education referral. Both items are part of the general education
practices section and are critical elements to ensuring an effective
process for making referrals to special education. Having ineffective
general education practices might create a bottleneck in the system
where students are made eligible for special education who otherwise
would not be. ED&D respondents also indicated lower favorable responses
with regard to having data routines and implementing routines well as
part of their multi-tiered system of support. An effective system of
support ensures students' needs are aligned to the right tiers of
instruction. Failure to implement an effective tiered instructional
system leads to faulty decisions about how students respond to
instruction. Finally, ED&D participants indicated lower preferred
ratings for gathering data on ethnic and racial breakdowns of children
being referred to special education and/or suspension or expulsion.
These questions emerge from the disproportionality section of the
survey. Knowledge about student demographics is an important aspectin
the decision-making process for special education eligibility and should
be better understood by the districts ED&D supports. Therefore, these
organizations supported by ED&D might benefit from additional resources
to improve these weaker areas so they are more situated to reduce
disproportionality in the future.
